#Principal Components Analysis (PCA)

#Load Data
```{r}
data(USArrests)
data = USArrests
states = row.names(data) # extract the state names
variable = names(data) # extract the variable names
head(data)
```

# Data Standerdization
```{r}
for (i in 1:ncol(data)){
  data[,i]=(data[,i]-mean(data[,i]))/sqrt(var(data[,i]))
}
head(data)
plot(data)
```

# PCA using prcomp()
```{r}
pr.out = prcomp(data, scale=FALSE)
print(pr.out)

# show the z-matrix:
pr.out$x[,1:2]
plot(data.frame(pr.out$x))

# show the first 2 columns of the Z matrix (if q=2)
pr.out$x[,c(1:2)]

plot(data.frame(pr.out$x[,c(1:2)]))

# biplot:
biplot(pr.out,scale=0)
```
1st frame: of the 4x4 matrix PC1 = column 1 of Phi,..., PC2 = col 2 of phi 2 and so on. We can choose how many columns to take though, in this case 2.
4th frame: shows how visualize the data in 2d.


# PCA 3D visualization
```{r}
library(rgl) # package for 3D visualization

plotPCA <- function(x, nGroup, text) {  # user-defined function
    n <- ncol(x) 
    if(!(n %in% c(2,3))) { # check if 2d or 3d
        stop("x must have either 2 or 3 columns")
    }

    fit <- hclust(dist(x), method="complete") # cluster
    groups <- cutree(fit, k=nGroup)

    if(n == 3) { # 3d plot
        plot3d(x, col=groups, type="s", size=1, axes=F)
        axes3d(edges=c("x--", "y--", "z"), lwd=3, axes.len=2, labels=FALSE)
        text3d(x,texts=text)
        grid3d("x")
        grid3d("y")
        grid3d("z")
    } else { # 2d plot
        maxes <- apply(abs(x), 2, max)
        rangeX <- c(-maxes[1], maxes[1])
        rangeY <- c(-maxes[2], maxes[2])
        plot(x, col=groups, pch=19, xlab=colnames(x)[1], ylab=colnames(x)[2], xlim=rangeX, ylim=rangeY)
        lines(c(0,0), rangeX*2)
        lines(rangeY*2, c(0,0))
    }
}

plotPCA(pr.out$x[,1:3], 4, text=rownames(data))
```

#A Simulation Study 1
```{r}
set.seed(100)
x1 = rnorm(50,0,1)
x2 = rnorm(50,x1,2)
x3 = rnorm(50,x2,1)
x4 = rnorm(50,x3,0.5)
x5 = rnorm(50,x4,1)
x6 = rnorm(50,0,1)
x7 = rnorm(50,0,1)
x8 = rnorm(50,0,1)
x9 = rnorm(50,0,1)
x10 = rnorm(50,0,1)
data = cbind(x1, x2, x3, x4, x5,
			x6, x7, x8, x9, x10)
rownames(data) = paste("A",seq(1,50,1),sep="-")

# data standardization
for (i in 1:ncol(data)){
  data[,i]=(data[,i]-mean(data[,i]))/sqrt(var(data[,i]))
}

# PCA using prcomp()
pr.out = prcomp(data, scale=FALSE)
biplot(pr.out,scale=0,var.axes = TRUE)
```

#A Simulation Study 2
```{r}
set.seed(100)
x1 = rnorm(50,0,1)
x2 = rnorm(50,x1,2)
x3 = rnorm(50,x2,1)
x4 = rnorm(50,x3,0.5)
x5 = rnorm(50,x4,1)
x6 = rnorm(50,0,1)
x7 = rnorm(50,0,1)
x8 = rnorm(50,0,1)
x9 = rnorm(50,0,1)
x10 = rnorm(50,0,1)
data = cbind(x1, x2, x3, x4, x5,
			x6, x7, x8, x9, x10)
rownames(data) = paste("A",seq(1,50,1),sep="-")

data[46,] = data[45,]+rnorm(10,0,0.1)
data[47,] = data[46,]+rnorm(10,0,0.1)
data[48,] = data[47,]+rnorm(10,0,0.1)
data[49,] = data[48,]+rnorm(10,0,0.1)
data[50,] = data[49,]+rnorm(10,0,0.1)

# data standardization
for (i in 1:ncol(data)){
  data[,i]=(data[,i]-mean(data[,i]))/sqrt(var(data[,i]))
}

# PCA using prcomp()
pr.out = prcomp(data, scale=FALSE)
biplot(pr.out,scale=0,var.axes = TRUE)

plotPCA(pr.out$x[,1:3], 3, text=rownames(data))
```

#PVE, proportion of variance explained of the z is 
```{r}
data(USArrests)
data = USArrests
for (i in 1:ncol(data)){
  data[,i]=(data[,i]-mean(data[,i]))/sqrt(var(data[,i]))
}
pr.out = prcomp(data, scale=FALSE)

PVE = pr.out$sdev^2/sum(pr.out$sdev^2)
barplot(PVE, names=paste("Z", 1:4, sep=" "))
```

#PC regression on the Prostate Cancer Data
```{r}
load("Prostate.RData")
data = Prostate
X = data[,1:8]

# data standardization
for (i in 1:ncol(X)){
  X[,i]=(X[,i]-mean(X[,i]))/sqrt(var(X[,i]))
}

# PCA on "data"
pr.out = prcomp(X, scale=FALSE)
pr.out
```










